Stack Safety as a Security Property

### 2020-07-01 main.tex
CP - 3

\begin{figure}
  \begin{minipage}{\textwidth}
    \begin{center}
\MemoryLabel{10em}{1.5em}{40}
\MemoryLabel{16em}{1.5em}{60}
\MemoryLabel{22em}{1.5em}{80}
$\cdots$
\memory{4}{gray}%
\memory{5}{green}[{\makebox[0pt]{Caller}}]%
\memory{5}{yellow}[{\makebox[0pt]{Args}}]%
\memory{8}{red}[{\makebox[0pt]{Callee}}]%
\memory{4}{gray}%
~$\cdots$\\
  \end{center}
  \end{minipage}\\
~\\
~\\
  \begin{minipage}{0.45\textwidth}
    \begin{center}
\begin{verbatim}
jal ...
lw r1 r0 40
\end{verbatim}
    \end{center}
  \end{minipage}~
  \begin{minipage}{0.45\textwidth}
    \begin{center}
\begin{verbatim}
jal ...
sw r0 r0 40
\end{verbatim}
    \end{center}
  \end{minipage}
  \caption{Examples of Unsafe Programs\bcp{The numbers don't line up with
      the boxes; the program fragments need comments explaining what each
      line does.}}
  \label{fig:stackunsafety}
\end{figure}

@@

The call stack is an increasingly popular target for attacks
exploiting memory safety vulnerabilities, with potential consequences
ranging from attackers reading or corrupting private stack data, to
highjacking a program's control flow. To prevent such attacks, many
techniques have been proposed in the literature\leo{cite, point out?
  stack canaries, cherry, nicks...}\bcp{I think pointing out makes a
  stronger point.} aiming to achieve ``stack
  safety''.

But what {\em is} this stack safety of which we speak? A precise formal
definition is lacking in the literature, with the notion usually being
defined negatively: through examples of stack unsafety
\rb{usually though not always; as noted below there are other
  approaches}\bcp{Yes, we have to be very careful about this}. Two classical
examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
\leo{Goal: on the left, a bad read; on the right, a bad write.}
\leo{Do we also need an example of ROP/control flow for the new
  part?}\bcp{If it is not too complicated to explain, we should definitely
  put it
  here.  If it is complicated, we can just say in text that there are other
  sorts of attacks too, and that we deal with them, and give a forward
  pointer.}
Recently, a \leo{inspiring/awesome/something good}\bcp{Maybe just say that
  we are only aware of one, quite recent, attempt to do something better} work reasoned
about a stack safety enforcement mechanism using capabilities\leo{cite};
but even here, stack safety is defined with respect to an abstract
machine that ``exhibits stack safety by construction,'' begging
the question of what {\em is} the property that holds by construction?

\leo{StackTokens} provides some intuition towards answering that question,
by informally describing two components that comprise stack safety:
local state encapsulation and well-bracketed control flow. The first
contribution of the present paper is to
turn this intuition into two formal \leo{extensional}\bcp{Not sure we can
  claim this} security properties,
that together form a formal definition of stack safety.
\rb{I think the merit of StkTokens goes beyond mere intuition. There is a lot of
rigor in it, even if some notions are implicit or not fully
developed.}\bcp{Yes, and let's remember that they are highly likely to be
reviewing this paper!  Let's be as generous as we can be without
under-selling our delta.}

The underlying notion behind local state encapsulation is the
intention to protect callers from callees: a callee should be not be
able to access or overwrite a caller's private data. We formalize this
intuition as a security property with both confidentiality and
integrity components.
%
Intuitively, {\em stack confidentiality} promises that a callee's
observable behavior is not affected by memory outside its stack frame
(e.g. a caller's private data or uninitialized memory).  {\em Stack
  integrity} guarantees that, if a function call ever returns, the caller's
stack data at the point of return will be identical to what it was before
the call.  \bcp{We need to point out somewhere that we are only dealing with
stack safety, though we discuss (somewhere) how our property can be combined
with the heap safety property of [Azevedo et al]}

While integrity and confidentiality are relatively standard security
properties, the stack-safety setting raises an interesting additional
challenge:
the notion of who the attacker is shifts throught execution. Not only
does a caller need to be protected from its callee, but the callee
itself needs to be proteced from {\em its} callees.
We address this challenge by varying the observation power
of the attacker at call points, and demonstrate that our policy is
implementable by pre-existing monitoring mechanisms. \bcp{The two parts of
  the sentence are about two very different things.  The second deserves its
  own paragraph, IMO.}

Similarly, the underlying notion behind well-bracketed control flow is
the need to protect the call structure itself: a callee should not be
allowed to jump to the middle of an arbitrary block of code, either
directly (via a jump instruction) or indirectly (by corrupting the return
address on the stack and then returning). \leo{There needs to be some
  informal description of the wbcf part here. Sean?}

In summary, we offer the following contributions:
\begin{itemize}
\item \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?} formal characterizations of the
  intuitive concepts of local state encapsulation and
  well-bracketed control flow~(\Cref{sec:lse-and-wbcf}).  As evidence that this
  property is strong enough to capture the folk intuition, we show
  that the stack unsafe examples from the literature also fail our
  property~(\Cref{sec:disc}).
\item As evidence that the property is not too strong, we show that it is
  realizible by
  a previously proposed micropolicy enforcement
  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
    substantial body of code}. We also present a variation of the
  property + policy that is lazy, easier/more efficient to implement,
  and provably weaker (\Cref{sec:lazy}).
\item A provably stronger (inductive? step-by-step/stepwise) variant
  of this property tuned for testing/verification.  We test the whole
  thing with a cool framework. \bcp{Which section?}
\end{itemize}
Section~\ref{sec:relwork} discusses related work, and
Section~\ref{sec:future} sketches directions for future work.
%
We begin with some technical preliminaries in
Section~\ref{sec:prelim}.

@@

% SNA: fixed
%\leo{PUT is completely ambiguous as a choice of abbreviation. Just ``Pre''
%  seems better}
The {\it prefUpTo} operation \(\PUT(\Trace,f)\) takes a trace \(T\)
and a predicate on elements \(f\) and gives the prefix of \(T\)
ending with the first element on which \(f\) holds, coinductively:

@@

\section{Stack Safety, Formally}
\label{sec:lse-and-wbcf}

In this section we will describe stack safety as a formal property. We
begin with our model of knowledge about the program structure\leo{I
  hate this}, and then dive in to local state encapsulation and
well-bracketed control flow.

@@

\leo{TODO: Figure out presentation}

\newcommand*{\add}{\text{add}}
\newcommand*{\addi}{\text{addi}}
\newcommand*{\sw}{\text{sw}}
\newcommand*{\lw}{\text{lw}}
\newcommand*{\jal}{\text{jal}}
\newcommand*{\jalr}{\text{jalr}}
%\begin{lstlisting}
%START:
%  addi r2 r2 12 ; allocate 3 words in the stack
%  sw r2 42 -8; store 42 in a private location
%  sw r2 0 -4; store 0 as an argument
%  sw r2 0 0; zero out the return location
%  jal r1 F ; call F and store return address
%
%F:
%  addi r2 r2 4 ; allocate for ra
%  sw r2 r1 -4 ; store the return address to sp-4
%  addi r2 r2 8 ; allocate for arg + ret
%  sw r2 17 -4; store 0 as an argument
%  sw r2 0 0; zero out the return location
%  jal r1 G ; call G and store return address
%
%G:
%  addi r2 r2 4 ; allocate for ra
%  sw r2 r1 -4 ; store the return address to sp-4
%
%\end{lstlisting}

Starting State:

Instruction memory in locations 0 - ..
PC at 100.
Stack memory starting at 1000.
Assume we're in a function.
Stack Pointer at 1040.
1040: 42.

\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{35.2em}{1.8em}{$\downarrow$}
\MemoryLabel{35em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{4}{gray}[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-18.7em}{0.6em}{42}
\MemoryLabel{-16.4em}{0.6em}{0}
\\
\end{center}

Program:

\begin{verbatim}
top{
  return f(0);
}

f(x){
  return (x + g(17));
}

g(x){
  return (x + 1);
}
\end{verbatim}

\newcommand*{\instr}[3]{ & #1: & #2 & ; \text{\em #3}\\}
\newcommand*{\clabel}[1]{#1: & & &\\}
\newcommand*{\rsp}{sp}
\newcommand*{\rra}{ra}
\[
\begin{array}{llll}
  \clabel{TOP}
  \instr{100}{\addi ~ \rsp ~ \rsp ~ 8}{allocate 2 words in stack}
  \instr{104}{\sw ~ \rsp ~ r_0 ~ -4}{store 0 as an argument}
  \instr{108}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
  \instr{112}{\jal ~ \rra ~ F}{call F and store the return address}
  & & & \\
  & & & \\
  \clabel{F}
  \instr{140}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
  \instr{144}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
  \instr{148}{\addi ~ \rsp \rsp ~ 8}{allocate for nested arg + ret}
  \instr{152}{\addi ~ r_4 ~ r_0 ~ 17}{put 17 in register $r_4$}
  \instr{156}{\sw ~ \rsp ~ r_4 ~ -4}{store 17 as the argument to G}
  \instr{160}{\sw ~ \rsp ~ r_0 ~ 0}{zero out the return location}
  \instr{164}{\jal ~ \rra ~ G}{call G and store the return address}
  \instr{168}{\lw ~ r_4 ~ \rsp}{load the return value of G to $r_4$}
  \instr{172}{\lw ~ r_5 ~ \rsp -12}{load the argument of F to $r_5$}
  \instr{176}{\add ~ r_4 ~ r_4 ~ r_5}{add $r_4$ and $r_5$}
  \instr{172}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as ther return of F}
  \instr{176}{\addi ~ \rsp ~ \rsp ~ -8}{deallocate local state}
  \instr{180}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
  \instr{184}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
  \instr{188}{\jalr ~ \rra ~ \rra ~ 0}{return to TOP}
  & & & \\
  & & & \\
  \clabel{G}
  \instr{200}{\addi ~ \rsp ~ \rsp ~ 4}{allocate for the return address}
  \instr{204}{\sw ~ \rsp ~ \rra ~ 0}{store the return address to sp}
  \instr{208}{\lw ~ r_4 ~ \rsp ~ -12}{load the argument of G to $r_4$}
  \instr{212}{\addi ~ r_4 ~ r_4 ~ 1}{increment $r_4$ by 1}
  \instr{216}{\sw ~ \rsp ~ r_4 ~ -8}{store $r_4$ as the return value}
  \instr{220}{\lw ~ \rra ~ \rsp ~ -4}{load the return address from the stack pointer}
  \instr{224}{\addi ~ \rsp ~ \rsp ~ -4}{restore the stack pointer of F}
  \instr{228}{\jalr ~ \rra ~ \rra ~ 0}{return to F}

%  addi r2 r2 8 ; allocate for arg + ret
%  sw r2 17 -4; store 0 as an argument
%  sw r2 0 0; zero out the return location
%  jal r1 G ; call G and store return address

\end{array}
\]

>>>>

\begin{figure}
  \begin{minipage}{\textwidth}
    \begin{center}
\MemoryLabel{10em}{1.5em}{40}
\MemoryLabel{16em}{1.5em}{60}
\MemoryLabel{22em}{1.5em}{80}
$\cdots$
\memory{4}{gray}%
\memory{5}{green}[{\makebox[0pt]{Caller}}]%
\memory{5}{yellow}[{\makebox[0pt]{Args}}]%
\memory{8}{red}[{\makebox[0pt]{Callee}}]%
\memory{4}{gray}%
~$\cdots$\\
  \end{center}
  \end{minipage}\\
~\\
~\\
  \begin{minipage}{0.45\textwidth}
    \begin{center}
\begin{verbatim}
jal ...
lw r1 r0 40
\end{verbatim}
    \end{center}
  \end{minipage}~
  \begin{minipage}{0.45\textwidth}
    \begin{center}
\begin{verbatim}
jal ...
sw r0 r0 40
\end{verbatim}
    \end{center}
  \end{minipage}
  \caption{Examples of Unsafe Programs}
  \label{fig:stackunsafety}
\end{figure}

@@

The call stack is an increasingly popular target for attacks
exploiting memory safety vulnerabilities, with potential consequences
ranging from attackers reading or corrupting private stack data, to
highjacking a program's control flow. To prevent such attacks, many
techniques have been proposed in the literature\leo{cite, point out?
  stack canaries, cherry, nicks...} aiming to achieve {\em stack
  safety}.

But what {\em is} stack safety? A precise formal definition of stack
safety is lacking from the literature, with the notion only being
defined negatively: through examples of stack unsafety
\rb{usually though not always; as noted below there are other approaches}. Two classical
examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
\leo{Goal: on the left, a bad read; on the right, a bad write.}
\leo{Do we also need an example of ROP/control flow for the new part?}
Recently, a \leo{inspiring/awesome/something good} work reasoned
about a stack safety enforcement mechanism using capabilities\leo{cite};
while impressive, stack safety was defined with respect to an abstract
machine that ``exhibits stack safety by construction''. Which begs
the question, what {\em is} the property that holds by construction?

\leo{StackTokens} provides some intuition towards answering that question,
by informally describing two components that comprise stack safety:
local state encapsulation and well-bracketed control flow. In this paper
we turn this intuition into two formal \leo{extensional} security properties,
that together form a formal definition of stack safety.
\rb{I think the merit of StkTokens goes beyond mere intuition. There is a lot of
rigor in it, even if some notions are implicit or not fully developed.}

The underlying notion behind local state encapsulation is the
intention to protect callers from callees: a callee should be not be
able to access or overwrite a caller's private data. We formalize this
intuition as a security property, with a confidentiality and an
integrity component.
%
At a high level, {\em stack confidentiality} ensures that a callee's
observable behavior is not affected by memory outside its stack frame
(e.g. a caller's private data or uninitialized memory).  {\em Stack
  integrity} guarantees that if a function call returns, the caller's
stack data will be identical to what it was before the call.

While integrity and confidentiality are relatively standard security
properties, this setting provides an interesting additional challenge:
the notion of who the attacker is shifts throught execution. Not only
does a caller need to be protected from its callee, but the callee
itself needs to be potentially proteced from {\em its} callees.  In
this paper we address this challenge by varying the observation power
of the attacker at call points, and demonstrate that our policy is
implementable by pre-existing monitoring mechanisms.

Similarly, the underlying notion behind well-bracketed control flow is
the need to protect the call structure itself: a callee should not be
allowed to jump to the middle of an arbitrary block of code either
directly (via a jump instruction) or indirectly (by corrupting stack
data and returning). \leo{There needs to be some informal description
  of the wbcf part here. Sean?}

We begin by giving some background in Section~\ref{sec:prelim}.
Our contributions are:
\begin{itemize}
\item \leo{More? Extensional} formal characterizations of the
  intuitive concepts of local state encapsulation~\ref{sec:lse} and
  well-bracketed control flow~\ref{sec:wbcf}.  Evidence that this
  property is strong enough to capture the folk intuition: we show
  that the stack unsafe examples from the literature also fail our
  property.
\item Evidence that the property is not too strong. It's realizible by
  a previously proposed micropolicy enforcement
  mechanism~\ref{sec:micropolicy}. \leo{And it says yes on a
    substantial body of code} We then present a variation of the
  property + policy that is lazy, easier/more efficient to implement,
  and provably weaker.
\item A provably stronger (inductive? step-by-step/stepwise) variant
  of this property tuned for testing/verification.  We test the whole
  thing with a cool framework.
\end{itemize}
Finally, in Section~\ref{sec:relwork} we discuss related work and
sketch directions for future work in Section~\ref{sec:future}.

@@

\leo{PUT is completely ambiguous as a choice of abbreviation. Just ``Pre''
  seems better}
The {\it prefUpTo} operation \(\PUT(\Trace,f)\) takes a trace \(T\)
and a predicate on elements \(f\) and gives the prefix of \(T\)
ending with the first element on which \(f\) holds, coinductively:

@@

\section{Stack Safety, Formally}

In this section we will describe stack safety as a formal property. We
begin with introducing a running example that will help keep things
concrete throughout this section, continue with our model of
\leo{compiler-provided?} knowledge about the program structure, and
then dive in to formal definitions local state encapsulation and
well-bracketed control flow.

\begin{figure}[b]
  \begin{centering}
    \begin{minipage}{0.5\textwidth}
\begin{verbatim}
f{
  int x, y;
  ...
  y = g(0);
  ...
}

g(x){
  return (x + h(17));
}

h(x){
  return (x + 1);
}
\end{verbatim}
\end{minipage}
\end{centering}
\caption{High-level Program with Simple Nested Calls}
\label{fig:running-program}
\end{figure}

@@

Our running example consists of three functions $f$, $g$, and $h$ that
perform a nested sequence of calls. A high level C-like view of the
excerpt program can be found in Figure~\ref{fig:running-program}.
Suppose that the starting state for our running example is right
before $f$'s call to $g$. A rough example layout of the memory at that
point can be found in the top of Figure~\ref{fig:running-trace-b}:
\begin{itemize}
\item The low part of the memory (0-1000) is dedicated to instructions.
\item The stack starts at location 1000.
\item The stack of $f$ starts at 1040.
\item The location of $x$ is 1040 and stores $42$.
\item The location of $y$ is 1044 and stores $0$.
\end{itemize}
Directly below the layout, we summarize the parts of the register file
that are relevant to the example.

The rest of Figure~\ref{fig:running-trace-b} detail the execution of the
program, showing the machine state at call and return points, as well
as describing the full effect of each instruction in detail.
\leo{Unoptimized, etc.}

\newcommand*{\add}{\text{add}}
\newcommand*{\addi}{\text{addi}}
\newcommand*{\sw}{\text{sw}}
\newcommand*{\lw}{\text{lw}}
\newcommand*{\jal}{\text{jal}}
\newcommand*{\jalr}{\text{jalr}}
\newcommand*{\rsp}{\mathit{sp}}
\newcommand*{\rra}{\mathit{ra}}

\newcommand*{\tracerow}[5]{#1: & #2 & #3 & #4 & \text{#5}\\}

\begin{figure}

%Starting Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{35.2em}{1.8em}{$\downarrow$}
\MemoryLabel{35em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{8}{gray}[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-29.6em}{0.6em}{42}
\MemoryLabel{-27.4em}{0.6em}{0}
\\
\end{center}
% Starting Register File
\[
\begin{array}{cccccc}
  PC: 100 & r_0 : 0 & \rra : ? & \rsp : 1048 & r_4 : ? & r_5 : ? 
\end{array}
\]
\vspace*{0.2em}
\[
  \begin{array}{r|c|c|c|c}
    & PC & \mathit{Instruction} & \mathit{Effect} &  \\
    \hline
    \tracerow{1}{100}{\addi ~ \rsp ~ \rsp ~ 8}{\rsp \leftarrow 1056}{Allocate call frame}
    \tracerow{2}{104}{\sw ~ \rsp ~ r_0 ~ -8}{[1048] \leftarrow 0}{Store 0 as argument}
    \tracerow{3}{108}{\sw ~ \rsp ~ r_0 ~ -4}{[1052] \leftarrow 0}{Zero out result}
    \tracerow{4}{112}{\jal ~ \rra ~ 140}{PC \leftarrow 140; ~ \rra \leftarrow 116}{Make the call}
  \end{array}
  \]
~ \\
% At Call to G
% Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{41em}{1.8em}{$\downarrow$}
\MemoryLabel{40.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg + result}}]%
\memory{6}{gray}[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}
\MemoryLabel{-22.8em}{0.6em}{0}
\\
\end{center}
% Register File
\[
\begin{array}{cccccc}
  PC: 140 & r_0 : 0 & \rra : 116 & \rsp : 1056 & r_4 : ? & r_5 : ? 
\end{array}
\]
% G - Body - 1
\vspace*{0.2em}
\[
  \begin{array}{r|c|c|c|c}
    \tracerow{5}{140}{\addi ~ \rsp ~ \rsp ~ 4}{\rsp \leftarrow 1060}{Allocate for return address}
    \tracerow{6}{144}{\sw ~ \rsp ~ \rra ~ -4}{[1056] \leftarrow 116}{Store return address}
    \tracerow{7}{148}{\addi ~ \rsp ~ \rsp ~ 8}{\rsp \leftarrow 1068}{Allocate for call frame}
    \tracerow{8}{152}{\addi ~ r_4 ~ r_0 ~ 17}{r_4 \leftarrow 17}{Store 17 to $r_4$}
    \tracerow{9}{156}{\sw ~ \rsp ~ r_4 ~ -8}{[1060] \leftarrow 17}{Store 17 as the argument}
    \tracerow{10}{160}{\sw ~ \rsp ~ r_0 ~ -4}{[1064] \leftarrow 0 }{Zero out return}
    \tracerow{11}{164}{\jal ~ \rra ~ 200}{PC \leftarrow 200; ~ \rra \leftarrow 168}{Make the call to $h$}
  \end{array}
  \]
  ~ \\
% At Call to H
% Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{49em}{1.8em}{$\downarrow$}
\MemoryLabel{48.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
\memory{3}{gray}[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}
\MemoryLabel{-22.8em}{0.6em}{0}
\MemoryLabel{-20.6em}{0.6em}{116}
\MemoryLabel{-18.4em}{0.6em}{17}
\MemoryLabel{-16.2em}{0.6em}{0}
\\
\end{center}
% Register File
\[
\begin{array}{cccccc}
  PC: 200 & r_0 : 0 & \rra : 168 & \rsp : 1068 & r_4 : 17 & r_5 : ? 
\end{array}
\]
% H - Body
\vspace*{0.2em}
\[
\begin{array}{r|c|c|c|c}
  \tracerow{12}{200}{\addi ~ \rsp ~ \rsp ~ 4}{\rsp \leftarrow 1072}{Allocate for return address}
  \tracerow{13}{204}{\sw ~ \rsp ~ \rra ~ -4}{[1068] \leftarrow 168}{Store return address}
  \tracerow{14}{208}{\lw ~ r_4 ~ \rsp ~ -12}{r_4 \leftarrow 17}{Load the argument of $h$ to $r_4$}
  \tracerow{15}{212}{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{Increment $r_4$ by 1}\end{array}
\]
% At Middle of H
% Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{52em}{1.8em}{$\downarrow$}
\MemoryLabel{51.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
\memory{1}{cyan}%
\memory{2}{gray}[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}
\MemoryLabel{-22.8em}{0.6em}{0}
\MemoryLabel{-20.6em}{0.6em}{116}
\MemoryLabel{-18.4em}{0.6em}{17}
\MemoryLabel{-16.2em}{0.6em}{0}
\MemoryLabel{-14.0em}{0.6em}{168}
\\
\end{center}
% Register File
\[
\begin{array}{cccccc}
  PC: 212 & r_0 : 0 & \rra : 168 & \rsp : 1072 & r_4 : 18 & r_5 : ? 
\end{array}
\]
\vspace*{0.2em}
\[
\begin{array}{r|c|c|c|c}
  \tracerow{16}{216}{\sw ~ \rsp ~ r_4 ~ -8}{[1064] \leftarrow 18}{Store $r_4$ as the return value}
  \tracerow{17}{220}{\lw ~ \rra ~ \rsp ~ -4}{\rra \leftarrow 168}{Load return address}
  \tracerow{18}{224}{\addi ~ \rsp ~ \rsp ~ -4}{\rsp \leftarrow 1068}{Restore the stack pointer of $g$}
  \tracerow{19}{228}{\jalr ~ \rra ~ \rra ~ 0}{PC \leftarrow 168}{Return to $g$}
  \end{array}
\]
\label{fig:running-trace-a}
\end{figure}

\begin{figure}
% After Return from H
% Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{49em}{1.8em}{$\downarrow$}
\MemoryLabel{48.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
\memory{3}{gray}
~$\cdots$
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}
\MemoryLabel{-22.8em}{0.6em}{0}
\MemoryLabel{-20.6em}{0.6em}{116}
\MemoryLabel{-18.4em}{0.6em}{17}
\MemoryLabel{-16.2em}{0.6em}{18}
\\
\end{center}
% Register File
\[
\begin{array}{cccccc}
  PC: 168 & r_0 : 0 & \rra : 168 & \rsp : 1068 & r_4 : 18 & r_5 : ? 
\end{array}
\]
% H - Body
\vspace*{0.2em}
\[
\begin{array}{r|c|c|c|c}
  \tracerow{20}{168}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
  \tracerow{21}{172}{\lw ~ r_5 ~ \rsp ~ -20}{r_5 \leftarrow 0}{Load the argument of $g$ to $r_5$}
  \tracerow{22}{176}{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow 18 + 0}{Add $r_4$ and $r_5$}
  \tracerow{23}{180}{\sw ~ \rsp ~ r_4 ~ -8}{[1056] \leftarrow 18}{Store $r_4$ as ther return of $g$}
  \tracerow{24}{184}{\addi ~ \rsp ~ \rsp ~ -8}{\rsp \leftarrow 1060}{Deallocate local state}
  \tracerow{25}{188}{\lw ~ \rra ~ \rsp ~ -4}{\rra \leftarrow 116}{Load return address}
  \tracerow{26}{192}{\addi ~ \rsp ~ \rsp ~ -4}{\rsp \leftarrow 1056}{Restore the stack pointer of $f$}
  \tracerow{27}{196}{\jalr ~ \rra ~ \rra ~ 0}{PC \leftarrow 116}{Return to $f$}
\end{array}
\]
% At Call to G
% Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{41em}{1.8em}{$\downarrow$}
\MemoryLabel{40.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg + result}}]%
\memory{6}{gray}
~$\cdots$
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}
\MemoryLabel{-22.8em}{0.6em}{18}
\\
\end{center}
% Register File
\[
\begin{array}{cccccc}
  PC: 116 & r_0 : 0 & \rra : 116 & \rsp : 1056 & r_4 : 18 & r_5 : 0
\end{array}
\]
\vspace*{0.2em}
\[
\begin{array}{r|c|c|c|c}
  \tracerow{28}{116}{\lw ~ r_4 ~ \rsp}{r_4 \leftarrow 18}{Load return value to $r_4$}
  \tracerow{29}{120}{\sw ~ \rsp ~ r_4 ~ -12}{[1048] \leftarrow 18}{Set $y$ to the return value}
  \tracerow{30}{124}{\addi ~ \rsp ~ \rsp ~ -8}{\rsp \leftarrow 1048}{Deallocate Call Frame}
\end{array}
\]
\vspace*{0.2em}  
% Ending Memory Layout
\begin{center}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{35.2em}{1.8em}{$\downarrow$}
\MemoryLabel{35em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
~$\cdots$
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
~$\cdots$
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{8}{gray}
~$\cdots$
\MemoryLabel{-29.6em}{0.6em}{42}
\MemoryLabel{-27.4em}{0.6em}{0}
\\
\end{center}
% Starting Register File
\[
\begin{array}{cccccc}
  PC: 128 & r_0 : 0 & \rra : 116 & \rsp : 1048 & r_4 : 18 & r_5 : 0
\end{array}
\]
\caption{Detailed Execution Trace}
\label{fig:running-trace-b}
\end{figure}
